---
title: "Obesity"
author: "Darwin Stahlback , Ani Lamichhane, Karen Vuong"
date: "12/15/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mosaic)
library(mosaicData)
library(lmtest)
library(skimr)
library(broom)
library(readr)
library(dplyr)
library(janitor)
library(dplyr)
```

```{r obesity}
#import data
obesity <- read_csv("Group-M-Obesity-Data.csv")
```

### Data wrangling:

```{r}
obesity <- obesity %>%  clean_names() #all column names to lower case
obesity$caec <- tolower(obesity$caec) #all column values to lower case
obesity$mtrans <- tolower(obesity$mtrans)#all column values to lower case
obesity$n_obeyesdad  <- tolower(obesity$n_obeyesdad)#all column values to lower case



obesity <- obesity %>%
  mutate(massbodyindex = obesity$weight/(obesity$height*obesity$height)) #this takes original n_obeyesdad categoricals back to numeric
obesity$is_obese = ifelse(obesity$massbodyindex<30,0,1) # creates binary if obese 0 = no / 1 = yes
is.num <- sapply(obesity, is.numeric)
obesity[is.num] <- lapply(obesity[is.num], round, 3) # format all numeric column to three places

```

```{r}
#renaming variables
colnames(obesity)[colnames(obesity) == "family_history_with_overweight"] = "family_history"

colnames(obesity)[colnames(obesity) == "favc"] = "food_between_meals"

colnames(obesity)[colnames(obesity) == "caec"] = "main_meals"

colnames(obesity)[colnames(obesity) == "scc"] = "monitor_calories"

colnames(obesity)[colnames(obesity) == "calc"] = "alcohol"

colnames(obesity)[colnames(obesity) == "mtrans"] = "transportation"

names(obesity)

obesityuse = obesity%>%
  select(gender, age, height, weight, family_history, food_between_meals, main_meals, smoke, monitor_calories, alcohol, transportation, is_obese)

```

### Variable Selection Multiple Logistic Regression:

```{r}
logmvarselcm1 = glm(is_obese ~ gender+age+height+family_history+food_between_meals+main_meals+smoke+monitor_calories+alcohol+transportation, data = obesityuse, family = binomial)
summary(logmvarselcm1)
```

```{r}
logmvarselcm2 = glm(is_obese ~ gender+age+height+family_history+food_between_meals+main_meals+smoke+monitor_calories+transportation, data = obesityuse, family = binomial)
summary(logmvarselcm2)
```

```{r}
logmvarselcm3 = glm(is_obese ~ gender+age+height+family_history+food_between_meals+main_meals+monitor_calories+transportation, data = obesityuse, family = binomial)
summary(logmvarselcm3)
```

### Nested likeyhood test:

model without transportation and main_meals:
```{r}
logmvarselcm3.1 = glm(is_obese ~ gender+age+height+family_history+food_between_meals+monitor_calories, data = obesityuse, family = binomial)
```

model without transportation:
```{r}
logmvarselcm3.2 = glm(is_obese ~ gender+age+height+family_history+food_between_meals+main_meals+monitor_calories, data = obesityuse, family = binomial)
```

model without main_meals:
```{r}
logmvarselcm3.3 = glm(is_obese ~ gender+age+height+family_history+food_between_meals+monitor_calories+transportation, data = obesityuse, family = binomial)
```

```{r}
lrtest(logmvarselcm3, logmvarselcm3.3)
```
```{r}
lrtest(logmvarselcm3.3, logmvarselcm3)
```

### Model and Interpretation:

```{r}
bestmodel = 
summary(bestmodel)
exp(coef(bestmodel))
```
Odds Space:

(copy and paste the sentences on the presentation)

### Condition:

Independence: Does not violate independence because the rows do not include time or spatial units.

Randomness: Obesity is a random factor. You don't really know if someone will be obese or not. 

Linearity of the Logit: 

```{r}
slogweight1 = glm(is_obese ~ weight, data = obesityuse, family = binomial)

slogweight <- augment(slogweight1, data = obesityuse)
slogweight <- slogweight %>%
  mutate(odds = exp(.fitted),
         probability = odds / (1 + odds))
```
```{r}
ggplot(slogweight, aes(x = weight)) +
  geom_line(aes(y = .fitted))
```
```{r}
obesityuse <- obesityuse %>%
  mutate(weightGroup = cut(weight, breaks = 10))
obesityuse %>%
  group_by(weightGroup) %>%
  skim_without_charts(is_obese)
```
```{r}
obesityuse_binned <- obesityuse %>%
  group_by(weightGroup) %>% 
  summarize(binnedis_obese = mean(is_obese), binnedweight = mean(weight)) %>%
  mutate(logit = log(binnedis_obese/(1-binnedis_obese)))
```
```{r}
ggplot(obesityuse_binned) +
  geom_point(aes(x = binnedweight, y = logit)) +
  geom_line(data = slogweight, aes(x = weight, y = .fitted))
```

```{r}
slogage1 = glm(is_obese ~ age, data = obesityuse, family = binomial)

slogage <- augment(slogage1, data = obesityuse)
slogage <- slogage %>%
  mutate(odds = exp(.fitted),
         probability = odds / (1 + odds))
```
```{r}
ggplot(slogage, aes(x = age)) +
  geom_line(aes(y = .fitted))
```
```{r}
obesityuse <- obesityuse %>%
  mutate(ageGroup = cut(age, breaks = 10))
obesityuse %>%
  group_by(ageGroup) %>%
  skim_without_charts(is_obese)
```
```{r}
obesityuse_binned <- obesityuse %>%
  group_by(ageGroup) %>% 
  summarize(binnedis_obese = mean(is_obese), binnedage = mean(age)) %>%
  mutate(logit = log(binnedis_obese/(1-binnedis_obese)))
```
```{r}
ggplot(obesityuse_binned) +
  geom_point(aes(x = binnedage, y = logit)) +
  geom_line(data = slogage, aes(x = age, y = .fitted))
```

```{r}
slogheight1 = glm(is_obese ~ height, data = obesityuse, family = binomial)

slogheight <- augment(slogheight1, data = obesityuse)
slogheight <- slogheight %>%
  mutate(odds = exp(.fitted),
         probability = odds / (1 + odds))
```
```{r}
ggplot(slogheight, aes(x = height)) +
  geom_line(aes(y = .fitted))
```
```{r}
obesityuse <- obesityuse %>%
  mutate(heightGroup = cut(height, breaks = 10))
obesityuse %>%
  group_by(heightGroup) %>%
  skim_without_charts(is_obese)
```
```{r}
obesityuse_binned <- obesityuse %>%
  group_by(heightGroup) %>% 
  summarize(binnedis_obese = mean(is_obese), binnedheight = mean(height)) %>%
  mutate(logit = log(binnedis_obese/(1-binnedis_obese)))
```
```{r}
ggplot(obesityuse_binned) +
  geom_point(aes(x = binnedheight, y = logit)) +
  geom_line(data = slogheight, aes(x = height, y = .fitted))
```


### Probability Space:
```{r, fig.width = 10}
mlogtransportation = glm(is_obese ~ age + transportation, data = obesityuse, family = binomial)

mlogtransportation = augment(mlogtransportation, data = obesityuse) %>%
  mutate(odds = exp(.fitted), 
         probability = odds / (1 + odds))

ggplot(mlogtransportation, aes(x = age, color = transportation)) +
  geom_point(aes(y = is_obese)) +
  geom_line(aes(y = probability))
```

```{r, fig.width = 10}
mlogfood_between_meals = glm(is_obese ~ age + food_between_meals, data = obesityuse, family = binomial)

mlogfood_between_meals = augment(mlogfood_between_meals, data = obesityuse) %>%
  mutate(odds = exp(.fitted), 
         probability = odds / (1 + odds))

ggplot(mlogfood_between_meals, aes(x = age, color = food_between_meals)) +
  geom_point(aes(y = is_obese)) +
  geom_line(aes(y = probability))
```

```{r, fig.width = 10}
mloggender = glm(is_obese ~ age + gender, data = obesityuse, family = binomial)

mloggender = augment(mloggender, data = obesityuse) %>%
  mutate(odds = exp(.fitted), 
         probability = odds / (1 + odds))

ggplot(mloggender, aes(x = age, color = gender)) +
  geom_point(aes(y = is_obese)) +
  geom_line(aes(y = probability))
```

```{r, fig.width = 10}
mlogmain_meals = glm(is_obese ~ age + main_meals, data = obesityuse, family = binomial)

mlogmain_meals = augment(mlogmain_meals, data = obesityuse) %>%
  mutate(odds = exp(.fitted), 
         probability = odds / (1 + odds))

ggplot(mlogmain_meals, aes(x = age, color = main_meals)) +
  geom_point(aes(y = is_obese)) +
  geom_line(aes(y = probability))
```

```{r, fig.width = 10}
mlogmonitor_calories = glm(is_obese ~ age + monitor_calories, data = obesityuse, family = binomial)

mlogmonitor_calories = augment(mlogmonitor_calories, data = obesityuse) %>%
  mutate(odds = exp(.fitted), 
         probability = odds / (1 + odds))

ggplot(mlogmonitor_calories, aes(x = age, color = monitor_calories)) +
  geom_point(aes(y = is_obese)) +
  geom_line(aes(y = probability))
```
```{r, fig.width = 10}
mlogfamily_history = glm(is_obese ~ age + family_history, data = obesityuse, family = binomial)

mlogfamily_history = augment(mlogfamily_history, data = obesityuse) %>%
  mutate(odds = exp(.fitted), 
         probability = odds / (1 + odds))

ggplot(mlogfamily_history, aes(x = age, color = family_history)) +
  geom_point(aes(y = is_obese)) +
  geom_line(aes(y = probability))
```

### Model Compared to Mean and Varinace of our Model:

```{r, fig.width = 10}
myplot = ggplot(obesityuse, aes(x = age, y = is_obese)) + geom_point(alpha = 0.5)
avg = mean(obesityuse$is_obese)
avg
myplot + geom_hline(yintercept = avg)
```

0.4613927 percent of people are obese. 

Our model is better than the mean because it explains % of the variability.

